### 2017-11-29 testdocs/test.tex
whatever

\documentclass{article}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage[margin=1.15in]{geometry}
\author{aodq}

\lstset{frame=tb,
  basicstyle=\footnotesize,
  language=C,
  aboveskip=2mm,
  belowskip=2mm,
  frame=none,
  keepspaces=true,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  breaklines=false,
  breakatwhitespace=false,
  tabsize=2
}

\lstset{language=C}

\title{DTOADQ}

\begin{document}
  \maketitle
  \section{Introduction}

  \section{Signed Distance Fields}
  Signed Distance Fields (SDFs) are a mathematical tool to represent a 3D
primitive by describing its surface via a distance field. Given an arbitrary
point in space, the SDF of a primitive is the distance to its nearest
point/surface. For example, the SDF of a sphere is $\Vert\:O\:\Vert - R$. With
SDFs, it's possible to march a ray along the scene until it hits the surface of
an SDF within a given bound, using the returned SDF distance as the distance to
march. Some example SDFs are shown below: \begin{align*}
  S(P, R) &= \Vert\:P\:\Vert - R & \text{Sphere with Radius}\\
  S(P, N) &= P \cdot N  & \text{Plane with Normal}\\
  S(P, H, R) &= <\Vert\:P_{x, z}\:\Vert - R, \: P_{y}> - H & \text{Torus with Radius and Hole}\\
  \end{align*}

  The biggest issue with SDFs is Lipschitz continuity. The SDFs themselves have
a distance gradient, that is, their bound. The bounds of a SDF is expected to be
at most 1 in order to remain Lipschitz continuous, as otherwise the SDF will
appear deformed as shown below.

  \includegraphics[]{lipschittzbound}

  This is a larger problem in Light Transport than it is in raytracing because
small discontinuities can impact the lighting of an entire scene. And to add to
that, most interesting operations (such as the meta balls shown in the image
above) will violate Lipschitz. The easiest way to fix this is to multiply the
distance field by a constant to stay below 1, but this is less of a fix and more
of a hack.

  In order to compute a normal of an SDF, you must perform a gradient
approximation: the central distance onf the SDF at P. For example, the
difference quotient:
  \[\frac{(f(x+h) - f(x))}{h}\]
  The most obvious way to calculate this is the six-point gradient; where for
each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
returns a vector pointing in the direction where the map's SDF changes the
most: the normal. A four point gradient is possible; as long as each each
point of the gradient is multiplied by $\epsilon$ before normalizing. The
speedup from six to four is worthwhile, as computing each gradient is
expensive as it involves a sdf map call. There are more utilities with
gradients not yet currently provided in DTOADQ but are worth mentioning, a
few are: an additional normal call using an alternative (higher-quality) SDF
map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
hand allows a cheap normal computation of primitives such as a spheres.
  \section{OpenCL}
	OpenCL, and GPU programming in general, is where I am least knowledgeable or
interested. Well, OpenCL is easy to work with, and I wrote my own OpenCL module
to handle OCL in a rather convenient method, for example, the DTOADQ kernel call
looks as such:
	\begin{lstlisting}
	shared_info.image.Lock(); // Acquire OCL Mem from OGL image object
	ocl.Run(
	  ocl.CLStoreMem(shared_info.rw_image),
	  shared_info.image.RWrite(),
	  ocl.CLStoreMem(shared_info.image_metadata),
	  shared_info.camera,
	  shared_info.timer,
	  ocl.CLPredefinedMem(imgs),
	  ocl.CLStoreMem(shared_info.ocl_material),
	  debug_values,
	  ocl.CLStoreMem(shared_info.rng_states),
	  // kernel size
	  shared_info.image.x, shared_info.image.y
	);
	shared_info.image.Unlock();

	\end{lstlisting}

	however the problem comes with dealing with the intrinsics of GPU; optimizing
code for the GPU and knowing how memory and threads/workers are handled.

  	One benefit of OpenCL is its ease to debug in comparison to OpenGL. One
immediate benefit is printf, which surprisingly even works on the GPU. To make
printf easily readable, I only use print on the center pixel of the image. Even
so, due to the complexity of light transport, the vast majority of time spent on
this project went directly towards debugging the DTOADQ kernel.

  	OpenCL also makes it relatively easy to send and receive data between the
host and GPU. In comparison to OpenGL, this is a big win, however, with
OpenGL4.3's Shader Storage Buffer Object, this is now no longer a selling point.

  	OpenCL also has huge drawbacks. The biggest problem is getting the results
to be rendered in OpenGL. TODO. Also, there are minor annoyances, like the lack
of a matrix data type, no function overloads, long compile times, or variadic
functions. There are also annoyances with GPU programming in general. No
dedicated random function. Though, possibly, long compile times is my fault for
using a mega-kernel instead of splitting it into many small kernels.

  	It seems that the GPU is not for me or DTOADQ and something I do not enjoy
to use nor am good at. However, for this project, I had to deal with it. To
overcome the lack of a random function, of which high quality uniform random
generation is necessary for Monte Carlo simulations, I used the MWC64X
algorithm. While there are better algorithms, they fail on branches which makes
it useless for DTOADQ. It seemed to work well enough. The downside is that each
worker/pixel needs its own random state, preserved across multiple kernel calls.
There are indefinitely better solutions.
  \section{Light Transport Introduction}
  \section{Bidirectional Path Tracing}
    The main foundation of the mathematics and theory have come from many
sources, such as Veach's thesis to PBR 3rd ed. and various open source path
tracing kernels such as NanoRT, Embree, etc. However, the implementation differs
from these in that the algorithm is optimized for the GPU. This section is about
transforming the algorithms and mathematical models into something easily and
efficiently implementable the GPU, however I will work from the ground up in the
theory and mathematics behind light transport.

    Calculating probability distribution functions is a very vital component in
Monte Carlo Sampling. The PDF states the probability of a random variable to
take the form of a given input. Its usefulness comes when a Monte Carlo
sample has a smaller or higher total contribution compared to other samples.
That is, one sample may have a higher probability distribution, and thus a
higher contribution to the path. That's the strategy behind Monte Carlo;
    \begin{align}
      \int f(x) \approx \frac{1}{N}\Sigma_0^N\frac{f(x)}{p(x)}
    \end{align}
    The PDF applies to everything from path contributions, solid angles of a
    BRDF, uniformly sampling a sphere, etc.

    The biggest conceptual barrier is that, for example, given a solid angle in
a hemisphere $\Omega$, the probability for a random uniform $X$ to equal
$\omega$ is 0. That's not what the PDF calculates, instead, it's the probability
that a random uniform $X$ could potentially equal $\omega$ when compared to
other $\omega\:$s.

    The good thing is that calculating these are relatively easy. The first step
is to find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
is the probability, $X$ is a uniformly random variable, and $x$ is a known
constant. This gives the probability that, for a formula $P$, where $X \in
\Omega$, and $\Omega$ is the set of all possible values for $CDF$, that $X \leq
x$. Then the PDF $p$ is defined by the derivative of CDF. To verify that the PDF
is correct, integrate it over $\Omega$, this final value should equal 1. This
makes sense intuitively -- it is a probability after all, so the sum of the
probabilities of every possible input must be 1. For an example, how I
calculated the PDF of an area light source
      $L^0_e(v) = L_o(P,\:\omega_o)$ 
    is quite simple. Firstly,the area light is defined as a sphere, and thus our
set of all possible values is correlated to surface area:
    $\Omega = [0,\:2.0f*\tau*R^2]$. Thus,
    \begin{align}
      CDF(x) = Pr(X \leq x) = \frac{x}{2.0*\tau*R^2}\\
      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0}{2.0*\tau*R^2}\\
      \int_0^{\Omega}p(x)dx = 1.0
    \end{align}
    In code
    \begin{lstlisting}
      float Light_PDF ( Emitter* m ) {
        return 1.0f/(2.0f*TAU*SQR(m->radius));
      }
    \end{lstlisting}

  Now, the importance of probability distribution functions comes when
performing Monte Carlo approximation on an integral. Take, for instance, the
most well-known form of the rendering equation:
  \begin{align}
    L_o(P, \omega_o) = L_e(P, \omega_o) + \int_{S}f_s(P, \omega_i,
    \omega_o) L_i(P, \omega_i) \vert cos\theta_i\vert d\omega_i
  \end{align}

  Where:
  \begin{list}{}{}
  	 \item $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.

  	 \item $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$
      (such as a lamp).

  	 \item $S$ is the unit sphere in which the angle $\omega_i$ is integrated
        over in respects to the normal at $P$. Also, this means, implicitly,
        that the equation is in respects to solid angle; the angle at which a
        primitive $A$ subtends $B$.

  	 \item $f_s$ is the Bidirectional Scattering Distribution Function. This
        will be detailed in the BSDF chapter, suffice it to say, it describes
        how photons are scattered (or absorbed) from point P in space from
        direction $\omega_i$ towards $\omega_o$ -- either through reflection
        (BRDF), or transmittance (BTDF). In volumetric path tracing, there is
        also subsurface scattering, but this phenomena is only approximated in
        DTQ.

  	 \item $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$

  	 \item And $cos\theta_i$ is the normal attenuation.
  \end{list}
  Conceptually, this is the behavior of photons bouncing off several surfaces of
a scene. While it's a very simple model, it's impossible to solve analytically
for anything but the most simple of scenes, mostly due to, for example, the
radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$. Another thing to
remember is that while it is shown to be integrated over a sphere, it is usually
shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.

  To compute this unbiasedly, it is required to use PDFs as described before to
approximate the integral using Monte Carlo simulation. The technique is to
generate $N$ uniformly random samples, weight them by their probability
distribution, and then average the results:

  \begin{align}
   \int{f(x)} \: \approx \frac{1}{N}\Sigma_0^N{\frac{f(x)}{p(x)}}
  \end{align}

  With standard "backward" path tracing, a photon would be emitted from a light
source and scatter in the scene until it hit an 'eye' or 'camera'. While this is
what happens in reality, in computer graphics, most of the samples never end up
hitting the camera/eye, so it is much more efficient to instead start from the
eye, and end at an emitter/light source. In this case, since we are integrating
over solid angle, the probably distribution function would be related to
sampling the BRDF. So for example, on a specular surface, the PDF is $1.0$ where
$\omega_o = reflect(\omega_i, N)$ and $0.0$ otherwise. An obvious optimization
is to only sample over the valid BSDF angles, which is known as importance
sampling. In DTQ, there are three BSDF functions:

  \begin{list}{}{}
  \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
  \item BSDF Sample: Samples a $\omega_o$ given $(P,\:\omega_i)$
  \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
  \end{list}

  There are many more optimizations to be made as well. For example, at each
scattering event, an additional photon can be sent towards each emitter. This is
what is known as next-event estimation. While this improves performance
drastically, naively averaging each contribution is incorrect, because each path
has a different probability to their paths/contributions. The technique to
sample them properly, multiple importance sampling, is related to BDPT, however
it will be detailed . The important thing to understand is that, in this
sampling strategy, there is a new PDF introduced which relates to the
probability of a path being generated, and this is critical to understand the
concept of BDPT.

  Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing.
Instead of starting from an eye or light, a 'photon' (path) is generated from
both, allowed to scatter in the scene for a few paths (the length is determined
by Russian Roulette), and connected. It's not exactly the end-points of the
paths that are being connected, but each vertex generated is connected to each
other. [SHOW EXAMPLE]. This actually has an interesting correlation to
next-event estimation, as it could be seen as 'extending' the amount of emitters
that can be connected to.

  For BDPT, like a lot of different light transport equations, things go much
more smoothly when the equation uses a different measure. In this case, the
measure is over surface area rather than solid angles. This looks like $v_0
\rightarrow v_1$. The transformation looks like so:
  \begin{align}
    f_s(v \rightarrow \dot{v} \rightarrow \ddot{v}) &= f_s(P, \omega_i,
                                                      \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e(P, \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v) L_e^1(v \rightarrow \dot{v})
  \end{align}
    That is, instead of looking at a point $P$ with an incoming and outgoing
    angle $(\omega_i,\:\omega_o)$, look at the path from $P_{-1}$ ($v$) to
    $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
    three different surfaces of the scene. There's a small change to be made in
the case of $L_e$, as DTOADQ [at least, for now], only has support for
area-light. Thus consider:

  \begin{align}
    L_e(P, \omega_o) &\equiv L_e(P) &\text{[ for area lights ]}\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v)&
  \end{align}

    The $cos\theta_i$ is no longer applicable, instead there is now a geometry
term, G, that converts a PDF with respect to solid angle to surface area, which
involves the inverse squared distance, and the cosine angle between the
geometric normals at $v$ and $\dot{v}$:

  \begin{align}
    cos\theta_n = v \cdot N(v) \\
    G(v \leftrightarrow \dot{v}) = V(v \rightarrow \dot{v})
        \frac{(cos\theta_n cos\dot{\theta_n})}{|| v - \dot{v}||^2}
  \end{align}

    Where $V$ is the visibility test. Thus the rendering equation with respects
    to area is:
  \begin{align}
    L(v \rightarrow \dot{v}) = L_e(v \rightarrow \dot{v}) \int_ML(v \rightarrow
    \dot{v}) G(v\leftrightarrow \dot{v}) f_s(v\rightarrow \dot{v} \rightarrow
    \ddot{v}) d_A
  \end{align}
    where $A$ is the area-measurement function on $M$, and $M$ is the set of all
points on the surface of a scene. That is, to say in regards to SDFs, $M$ is the
set of all points $P$ where $|map(P)| < \epsilon$. The area-measurement function
goes into measure theory, which while it is beyond my grasp currently to
compute, conceptually I understand a bit of it. The important thing is that $A$
is a function on the set of all subsets of $M$ that returns the measurement of
each set, and as long as the surfaces of the scene are sampled uniformly, this
never actually has to be computed. To back it up a little further, Veach
presented this as a means to integrate over a measure rather than using solid
angles. This proves that you can transform the rendering equation to different
measurements. For example, instead of integrating over points on surfaces, one
could integrate over an entire 3D space in order to perform volumetric path
tracing. Transforming the rendering equation using measurements would be
interesting in the future to look into.

    Anyways, with the three-point form now described, by recursively expanding,
so that the three point form is integrated over every possible set of possible
paths, this can be rewritten as the more usable format:
  \begin{align} I_j = \int_\Omega f_j(\bar{v}) d\mu(\bar{v}) \end{align}
    Where $\bar{v}$ is a path $z_0\rightarrow z_s \rightarrow y_t \rightarrow
    y_0 (eye \rightarrow light)$, $\Omega$ is the combination of all possible
    paths of any length, and $\mu$ is the area-product measure [the product of
all the expanded $A$]. Special care has to be taken with the definition of
    path itself, as in the source code, there never exists a path $\bar{v}$,
    only eye-path $\bar{z}$ and light-path . Monte Carlo approximation, as shown
above, is applied to the formula:
  \begin{align}
    I_j \approx \frac{1.0f}{N}\Sigma_0^NF\\
    F = \Sigma_s\Sigma_t\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}
  \end{align}

    where $s$ is the light-path length, and $t$ is the eye-path length. In
    DTOADQ, they are limited to $s+t \leq DEPTH$, a kernel-defined macro.
Finally, to transform $\bar{v}$ into two seperate paths, so that an eye and
light path may be generated, $\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}$ is
split into two with a connection edge $c$
  \begin{align}
    \frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})} \equiv
    \frac{f^L(\bar{y})}{p(\bar{y})} c(\bar{y}_s \leftrightarrow \bar{z}_t)
    \frac{f^E(\bar{z})}{p(\bar{z})}  
  \end{align}

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}

  $\alpha_{s|t}^{L|E}$ is the final radiance of either the emitter or eye path,
  and thus it is recursive in nature. That is, the value of $\alpha_i$ is
  dependent on $\alpha_{i-1}$, unless $i = 0$ in which case is the radiance of
  either the emitter or the eye path. However, this can be simplified further,
because of Helmhotz reciprocity, BRDFs return the same value if $wo$ and $wi$
are switched. This is also the reason why forward and backward path tracing
work.
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      1.0 &i = 0,\\
      \frac{
      L_e^0(V_0 \rightarrow V_1)}{p_A(V_1)} | 1.0 &i = 1,\\
      \frac{f_s(V_0 \rightarrow V_1 \rightarrow V_2)}{p_{\sigma}(V_0
      \rightarrow V_1)} &i > 1
    \end{cases}
  \end{align}
    For the case of $i = 0$, this would mean that the path does not exist. In
other words, forward or backward path tracing, the latter of which isn't
supported as camera lens are not defined in DTQ. For $i = 1$, it is either a
point on the surface of a light, or the origin point on the 'delta eye'. Lastly,
the BRDF is used to calculate the albedo of each primitive in the scene, and is
then divided by the PDF sigma, which is the probability density of $v_i
\rightarrow v_{i-1}$. $\alpha_1^L$ is shown below:

\begin{lstlisting}
  float3 N      = Sample_Cosine_Sphere(si, &pdf_pos),
         origin = light.origin + light.radius*N,
         dir    = Reorient_Angle(Sample_Cosine_Hemisphere(si, &pdf_dir), N);
  float pdf_pos = 1.0f/(2.0f*TAU*SQR(light.radius)),
        pdf_amt = 1.0f/EMITTER_AMT;
  pv->irradiance = light.emission/(pdf_pos*pdf_amt*pdf_fwd);
\end{lstlisting}

  Note that the irradiance is being calculated in this case, that is, the amount
  of flux arriving at point $pv$. Though in the case of $s=1$, this is more like
  the radiance. The irradiance at $t > 1$ and $s > 1$ is shown:

\begin{lstlisting}
  float3 sample_dir, sample_f;
  float sample_pdf;
  _BSDF_Sample(wi, N, vtx, &sample_dir, &sample_pdf, &sample_f, si);
  *irradiance = sample_f * (*irradiance) * fabs(dot(N, sample_dir))/sample_pdf;
\end{lstlisting}

  The connection strategy below describes how to connect the edges of the two
paths. There is no $t = 0$ case as no physical camera lens exists on the scene.
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(L_0 \rightarrow L_1), &s = 0, t > 1,\\
      f_s(L_0 \rightarrow L_1 \rightarrow L2) G(E_1 \leftrightarrow
      L_1) f_s(L_1 \rightarrow E_1 \rightarrow E_0), &s > 0, t > 1
    \end{cases}
  \end{align}
  In code:
  \begin{lstlisting}
Spectrum contribution = E1.irradiance * L1->irradiance;

// As there is no T=1 case, L1 -> E1 -> E0 [Le in s=1 case] must happen
contribution *= Subpath_Connection(L1, &E1, &E0);

// For s == 1, L1 = L0, a point on the surface of emitter, there is no
// L0 -> L1 -> E1 connection
if ( light_depth > 0 ) {
  // Connect L0 -> L1 -> E1
  contribution *= Subpath_Connection( L0,  L1, &E1);
}

// Geometric connection term [includes visibility check]
contribution *= Geometric_Term(&E1, L1, si, Tx);

  \end{lstlisting}

    Is the $s = 0, t > 1$ strategy worthwhile to compute? Those familiar with
next event estimation would know that there are situations in which this is
beneficial. Take, for example, a glossy surface with a small solid angle
subtending to the surface of a large light source. The $s=1$ sample strategy in
this case would only contribute when the $L0$ vertex is within that solid angle.
The effective strategy to sample in this case is $s = 0$. In code,
$\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
    \begin{lstlisting}
     float3 wi = normalize(E1.origin - E0.origin),
            wo = normalize(L1.origin - E1.origin);
     float3 albedo = _BRDF_F(wi, E1.normal, wo, E1.material, E1.mat_col);
     albedo /= BSDF_PDF(wo, wi, &E1);
     Spectrum contribution = E1.radiance * albedo;
     contribution *= Geometric_Term(&E1, &L1, si, Tx);
     contribution *= tlight.emission/Light_PDF(&tlight);
    \end{lstlisting}

    Another question to consider is if the path should be able to reflect on the
light source. While under real circumstances, this is allowed, the contribution
is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t =
0|1$ as the lens does not have a physical existence in the SDF map. The $t = 0$
case, where a photon from the light source hits the lens directly, is rather
pointless anyways, as a path generated by light-tracing would have to be able to
contribute to any pixel, which is not feasible on a GPU.
\\

The amount that each path contributes differ greatly, so naively weighing
each contribution uniformly by something like 1.0f, or even, by inverse path
length, will producy very noisy images, and would not be worth the additional
computation time that BDPT requires. Thus, as mentioned earlier, probability
distributions of the entire path are taken into account in order to generate
multiple importance samples.
    The formula for multiple importance sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
generated, while the denominator is the path density of other connection
strategies that could could have, in theory, created the path. The
straight forward implementation of this has a lot of problems gone unresolved in
Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses and describe these
problems. Primarily, $P_i(\bar{x})$ will overflow CLFloat (due to the distance
in the Geometric term), and the straight forward implementation is very long,
hard to debug, and has a poor time complexity. But the important part is the end
result of their solution:
  \begin{align}
    r_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = s,\\
      \frac{\overleftarrow{P}(\bar{x_i})}{\overrightarrow{P}(\bar{x_i})} *
      r_{i+1}(\bar{x}), & i < s,\\
      \frac{\overrightarrow{P}(\bar{x_{i-1}})}{\overleftarrow{P}(\bar{x_{i-1}})}
      * r_{i-1}(\bar{x}), & i > s
    \end{cases}
  \end{align}
    However, DTOADQ can simplify a bit further; unlike above, there are two
paths for which to evaluate. Specifically, in the above term, for where $i < s$,
we have $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$,
which is the light path (i is iterating towards light-length s). The eye
contribution is inversed from this, (also iterating towards s, hence why
$\bar{x}_{i-1}$ is used). . In DTOADQ's case, the weights are calculated in the
same direction relative to its originator; in this case the perspective of the
eye path. Not only this, but unlike any implementation I've found of BDPT, I
don't recalculate each path's probability and MIS every vertex. That is, usually
on calculating MIS for each connection, the MIS and probability values are
recalculated for each vertex in the path. As this is not possible given that
only the light-path exists in memory, an array for MIS/probability is stored
corresponding to each light vertex, and a single MIS/probability is stored
corresponding to $E_1$. Thus, the weight on each path looks like:
  \begin{align}
    \mathcal{W}_i(\bar{x}) \equiv r_i(\bar{x})\\
    \mathcal{W}_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = 0,\\
      \frac{\overleftarrow{p_i}(\bar{x_i})}{\overrightarrow{p_i}(\bar{x_i})} *
      \mathcal{W}_{i-1}(\bar{x}), & i > 0
    \end{cases}
  \end{align}
    And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like
  \begin{align}
    \mathcal{W}_i(\bar{x}) = \Pi_{n=2}^i(\frac{\overleftarrow
    {p_{\sigma}}(x_n)\overleftarrow {G}(x_n)}
    {\overrightarrow{p_{\sigma}}(x_n)\overrightarrow{G}(x_n)})
    * \frac{\overleftarrow{p_A}(x_1)}{\overrightarrow{p_A}(x_1)}
  \end{align}
    And now the MIS looks like:
  \begin{align}
    MIS_{s, t}(\bar{y}, \bar{z}) =
    \frac{1.0f}{\mathcal{W}_s(\bar{y}) + \mathcal{W}_t(\bar{z}) + 1.0f}
  \end{align}
    Thus, the final rendering equation for DTOADQ looks like such:
  \begin{align}
    P = \frac{1.0f}{N} \Sigma_{0}^{N} \Sigma_{s \le 0} \Sigma_{t \le 0}
    \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
  \end{align}

  \begin{align}
    P_i(\bar{x}_{s, t}) = P^L_s p^E_t
  \end{align}

  \begin{align}
    p^{L|E}_i(\bar{x}) =
    \begin{cases}
      1.0f, & i = 0,\\
      P_A(x_i), & i = 1,\\
      P_{\sigma}(x_{i-1} \rightarrow x_i) G(x_{i-1} \leftrightarrow x_i)
      P^{L|E}_{i-1}, & i > 1
    \end{cases}
  \end{align}

  \section{DTOADQ BRDF}
  DTOADQ supports a "universal" physically-based sampling and spectrum BRDF. The
spectrum BRDF is a combination of many different modified BRDF functions into
one. There is a microfacet, subsurface and diffusive component in the DTOADQ
spectrum BRDF, which are all combined together, with many variables to tune
these components. There is diffusive, glossy, specular and transmittive sampling
in the BRDF importance sample, with variables to allow specific control over
which strategies to choose and how they are combined together. Regardless of how
the BRDF is sampled, they use the same spectrum BRDF.

  In order to develop these BRDFs, I wrote an SDF model based off the mitsuba
model and wrote an IBL path-traced shader for quick prototyping;
shadertoy.com/view/XtSyDm . The main reason to not use DTOADQ was the lack of
cubemap support, which would have taken too long to implement.

  The Microfacet model consists of a fresnel, distribution and geometric term.
It's important to note that these have all been modified to use the half-vector
$H = \Vert\:L+V\:\Vert$. This gives better visuals and allows for a specific
energy conservation model. Anyways, the fresnel term approximates the fresnel
factor, it's based off both Schlick's approximation and Disney's diffusive
schlick fresnel, with modifications to use a $metallic$ term as well as an
albedo;
  \begin{lstlisting}
  const float3 Fresnel_L = Schlick_Fresnel(cos_NL),
               Fresnel_V = Schlick_Fresnel(cos_NV);
  const float F0 = m->fresnel * m->metallic,
  Fresnel_diffuse_90 = F0 * SQR(cos_HL);
  microfacet *= (1.0f - F0) * diffusive_albedo +
      mix(1.0f, Fresnel_diffuse_90, Fresnel_L) *
      mix(1.0f, Fresnel_diffuse_90, Fresnel_V);
  \end{lstlisting}
  \includegraphics[scale=0.20]{fresnel}

  The Geometric term is used to control how microfacets 'shadow' each other. The
DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector,
combined with an anisotropic term based off the GGX Anisotropic model:
  \begin{lstlisting}
  const float Param  = sqr(0.5f + sqr(m->roughness)),
  Aspect = sqrt(1.0f - m->anisotropic*0.9f),
  Ax     = Param/Aspect, Ay = Param*Aspect,
  GGX_NV = Smith_G_GGX_Correlated(cos_HL, cos_NV, Ax),
  GGX_HL = Smith_G_GGX_Correlated(cos_NV, cos_HL, Ay);
  microfacet *= 0.5f / (GGX_NV*Ax + GGX_HL*Ay);
  \end{lstlisting}
  \includegraphics[scale=0.20]{geometric}
  And finally, the Distribution term controls how microfacets are spread around
the normal. It's based off the Hyper Cauchy with a roughness and metallic
(shape) parameters [9]. The latter of which should most likely be its own
parameter in the future. The more readable mathematical model is shown:
  \begin{align}
  \frac{(M - 1)\:(\sqrt{2})^{2\:S\:-\:2}}
  	    {\pi\:R^{2}\:cos^{4}\theta\:(2 +
  	      \frac{tan^{2}\theta}{M^2})^M}
  \end{align}
  \includegraphics[scale=0.20]{distribution}

  These combine to form the microfacet under the half vector microfacet energy
conservation model [Edwards et al. 2006].
  \[
  	\frac{F*G*D}{4 \: (H \cdot V) \: ( (N \cdot L) \vee (N \cdot V) )}
  \]

  The diffusive component is a combination of both the subsurface and lambertian
component. The lambertian is simply $\alpha^{0.5} * \frac{1}{\pi}$, and the
subsurface component is based off the Henyey-Greenstein formula [Henyey],
modified for half-vec:
  \begin{align}
  	\frac{m(1 - r^2)}{4\pi}\:\frac{1}{(1 + r^2 - 2\:r\:
  		(H \cdot V))^{3/2}}
  \end{align}
  where m is metallic and r is roughness. This is then combined to the diffusive
component of the BRDF using Fresnel factor like so:
  \begin{lstlisting}
const float R = (-0.3f + 1.3f*m->roughness)*cos_HV,
            M = (1.0f - m->metallic)*5.0f;
const float Rr_term = M * (1.0f - sqr(R))*IPI *
                      (1.0f/(pow(1.0f + sqr(R) - 2.0*R*cos_HL, 3.0f/2.0f)));
const float3 Retro_reflection = diffusive_albedo * Rr_term *
                  (Fresnel_L + Fresnel_V + (Fresnel_L*Fresnel_V*(Rr_term)));
diffusive_albedo = mix(diffusive_albedo, Retro_reflection, m->subsurface);

  \end{lstlisting}
  TODO: With 0.95 to exaggerate

  The microfacet and diffusive components are then added together.

  \section{Video/Image Emitter}

  \section{Future Improvements}

  Every emitter is chosen uniformly, however a better system would weight the
  probability of each emitter by
  \[
    p_l = \frac{\phi_lr_l}{\Sigma_i\phi_ir_l}
  \]
  Thus emitters that have a higher contribution to the scene, that is, a larger
  emittance and radius, will have a higher probability to be selected.

        // What's defined: V1 -> V2 [we're evaluating V1] This is for a lens.
        // So, it's ultimately completely useless until an
        // actual lens exists in the SDF scene. Note that this isn't an actual
        // point on a surface of the scene (that's V2, next up for eval), this
        // is just an infinitesimal point based off where the camera currently
        // is, just like a delta light. So while the point actually exists on a
        // scene, the evaluation is on a "delta camera"

  \section{Bibliography}
  [9]
  Samuel D. Butler, Experimental and Theoretical Basis for a closed-Form Spectral BRDF model

  [Henyey]
   Diffuse Radation in the Galaxy, 1942. L, G. Henyey and J. L. Greenstein. 
\end{document}